---
title: "Motor Trend"
author: "MATIAS, V. A."
date: "30/03/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
<p> This relatory is the final project of the 
[Regression Models Course](https://www.coursera.org/learn/regression-models) 
offered by Johns Hopkings  University in the 
[Coursera Plataform](https://www.coursera.org/)

<p> The objective of the paper is review a collection of cars and explore 
the relationship that explains in a quantitative way the miles per gallon spent.
To do this, will be used strategies of Exploratory Data Analysis and Regression 
Models.

## Data
<p> The data set used can be retrieved in the base R software by the lines 
below:
```{r warning=FALSE, message=FALSE}
library(dplyr)

data("mtcars")
mtcars <- as_tibble(mtcars)

mtcars
```

<p> So we have 32 observations with 11 numeric attributes for each. Let's see 
the correlation between the attributes.
```{r warning=FALSE, message=FALSE}
library(corrplot)

corrplot(cor(mtcars), 
         method = "color",
         type = "upper", 
         addCoef.col = TRUE,
         diag = FALSE,
         number.cex = 0.7)
```
<p> The figure shows that mpg has a strong negative correlation for cyl, disp,
hp and wt (more intense red). Positive correlation can be found, mainly,
with drat, vs and am. This variables can be

## Trends


## Apendix
<p> mtcars attributes
