---
title: "Motor Trend"
author: "MATIAS, V. A."
date: "30/03/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
<p> This report is the final project of the 
[Regression Models Course](https://www.coursera.org/learn/regression-models) 
offered by Johns Hopkins  University in the 
[Coursera Plataform](https://www.coursera.org/). The paper's objective is to 
review a collection of cars and explore the relationship that explains 
quantitatively the miles per gallon spent. To do this will be used strategies 
of Exploratory Data Analysis and Regression Models.



## Trends: MPG and Transmission
<p> The data set used can be retrieved in the base R software by the command 
`data("mtcars")`. It has 32 observations with 11 numeric attributes for each. 
This paper analyzes the available data to understand if automatic transmissions 
are better than the manual for mpg and quantify the difference. The Conclusions 
section shows the answers. To know more about the data set and the variables 
use the command `?mtcars`.

### Simple Linear Regression
<p> Some exploratory analysis about the transmissions shows that automatic 
transmissions have lower mpg values (appendix, figure 1). To see if this 
relation follows a pattern, we'll do a simple linear regression with mpg 
(Miles/US gallon) as the outcome and am (transmission) as the explanatory 
variable.
```{r}
fit <- lm(mpg ~ factor(am), data = mtcars)
summary(fit)$coefficients
message("R Squared:", summary(fit)$r.squared, "\nAdjusted R Squared:", 
        summary(fit)$adj.r.squared)
```
<p> Is expected that the values of mpg for automatic transmission be around 
17.147 $\pm$ 1.125 and the cost to change the transmission for manual is equal 
to 7.245, getting around 24.392 $\pm$ 1.764 for manual transmission. The result 
of the t-test results in a P-value very low for transmission, which is good. 
The results of Multiple $R^2$ and Adjusted $R^2$, says that the linear model 
can explain around of one-third of the variation in the data, so _am_ alone 
brings some information about the mpg, but it's better to add more variables.

### Multivariable Linear Regression
<p> First, we need to find the main variables to put on a new model.
One way is using variables with significant correlation with mpg, but well 
uncorrelated between themselves. The correlation table can draw some insights 
to understand the pattern (appendix, figure 2), else can be done a model with 
all variables and, with him, extract that with lower p-values on the t-test.
```{r}
coefs <- summary(lm(mpg ~., mtcars))$coefficients
coefs[coefs[,4] < 0.35,4]
```
<p> As we can see, the most influenceable variables (lower p-values) are wt 
(Weight in 1000 lbs), qsec (1/4 mile time) and hp (Gross horsepower), plus am. 
Now we'll build one model with these four variables and another without hp. 
Then, the ANOVA is used to find the model with a lower p-value for F-test and 
get his results.
```{r}
mult_fit <- lm(mpg ~ factor(am) + wt + qsec, data = mtcars)
mult_fit_hp <- lm(mpg ~ factor(am) + wt + qsec + hp, data = mtcars)

p_values <- anova(fit, mult_fit, mult_fit_hp)$`Pr(>F)`[2:3]
message("P-Value without hp(", p_values[1], ")", 
        "\nis lower than the P-Value with hp(", p_values[2], ")")

summary(mult_fit)$coefficients
message("R Squared:", summary(mult_fit)$r.squared, "\nAdjusted R Squared:", 
        summary(mult_fit)$adj.r.squared)
```
<p> The results say that qsec and wt were added well in the model, with low 
p-values. The R-squared and adjusted R-square grows to approx. 0.85 and 0.83, 
respectively.


<p> The residuals (appendix, figure 3) showed no relevant problems and no 
relevant pattern compared with the fitted values. They yet tended to a linear 
tendency in normal QQ plot, although a small sinusoidal pattern can be detected,
it isn't anything impactful. The other residual plots show did not show an 
abnormal trend, having expected behavior.

## Conclusions
<p> As can be derived by the boxplot analysis of the data and the pattern 
**with other variables** showed before, automatic transmission tends to be 
better than manual transmissions for mpg. Analyzing the mpg trend just with 
the fact to be manual or automatic don't show information with significant 
confidence.

<p> So, mpg can be calculated as: mpg = 9.6175 + 2.9358$*$am - 
3.9165$*$wt + 1.2259$*$qsec, where am = 0 for automatic transmission and 1 for 
manual transmission. Following the trend of the chosen model, the automatic 
transmission starts with 9.6178 for mpg and increases/decreases with wt and 
qsec values. Manual transmissions, on another side, have a start increased in 
2.9358 mpg, so, starts with 12.5536 and changes the outcome by the values of qsec 
and wt.

## Appendix

```{r warning=FALSE, message=FALSE, fig.width=4, fig.height=4, fig.align='center',fig.cap="mpg X Transmission Boxplot"}
library(ggplot2)
library(dplyr)

mtcars$am <- if_else(mtcars$am == 0, true = "auto", false = "manual")

ggplot(mtcars) + 
  geom_boxplot(aes(x = am, y = mpg, group = am)) + 
  labs(x = "Transmission",
       y = "Milles per gallon (mpg)")
```

```{r warning=FALSE, message=FALSE, fig.width=8, fig.height=8, fig.align='center',fig.cap="Correlation Table"}
library(corrplot)
data("mtcars")

corrplot(cor(mtcars), 
         method = "color",
         type = "upper", 
         addCoef.col = TRUE,
         diag = FALSE,
         number.cex = 0.9)
```

```{r, fig.align='center',fig.cap="Residuals plots (multivariable model)"}
par(mfrow = c(2,2))
plot(mult_fit)
```