---
title: 'Statistical Inference'
subtitle: 'Central Limit Theorem review and some inference about teeth grow'
author: "MATIAS, V. A."
date: "05/04/2021"
output: pdf_document
---

# Overview
The first part of the report makes one thousand simulations of the exponential 
function to analyzes the Central Limit Theorem application under the means 
distribution. The second part is focused to applies statistic inference 
techniques bringing a better understanding of teeth growth.


# 1. Central Limit Theorem
The Central Limit Theorem (CLT) says that the averages of any distribution 
tend to the normal as the amount of data grows. The first topic examines 
this fact.

## 1.1 Simulation
We'll create a thousand exponential distributions and extract the means to 
build a new distribution that should be next to the normal. The exponential 
distribution is calculated as:

$f(x) = \lambda {e}^{- \lambda x}$

Where:

* $\lambda$ = mean/variance of the data, being that both variance and mean 
are the same in a Poisson Distribution

* _x_ = The numeric data related to the observations

In this report will be created 1000 of these distributions. Each of them has 
40 random values, and lambda equals 0.2, so the formula can be adapted to 
$f(x) = 0.2 {e}^{- 0.2 x}$. The code chunk below extracts the means of each of 
the exponential distributions, store them and plot the means distribution 
plot.

```{r, warning=FALSE, message=FALSE}
library(ggplot2)

mns <- c()
for (i in seq(1000)) {
  mns <- c(mns, mean(rexp(n = 40, rate = 0.2)))
}

ggplot() + 
  geom_density(aes(x = mns), col = "red") + 
  geom_density(col = 'blue',aes(rnorm(1000, mean = mean(mns), sd = sd(mns)))) 
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
testt <- t.test(mns, conf.level = 0.95)$conf.int
incon <- mean(mns) + c(1,-1) * qnorm(0.025) * sd(mns)/sqrt(length(mns))

length(mns[mns > testt[1] & mns < testt[2]])

```

## Sample Mean versus Theoretical Mean

## Sample Variance versus Theoretical Variance

## Distribution

# Teeth

## Data

## Inference

# Conclusions