---
title: "Statistical Inference"
author: "MATIAS, V. A."
date: "07/04/2021"
output: pdf_document
subtitle: 'Part 2: Basic Inferential Data Analysis about tooth growth '
---

# Overview 
The second part of the project analyzes the dataset ToothGrowth and does 
some exploratory data analysis, and applies basic statistical inference to 
bring relevant information

# Data
The dataset has 60 rows and 3 columns (appendix, chunk code 1). 
Each line has the length of the odontoblasts in the study 
(cell responsible for the growth of the teeth), the supplement used and the 
size of the dose (0.5, 1 or 2mg / day). The study was carried out on guinea 
pigs. Let's look at a random sample of 6 tuples.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library('dplyr')
data('ToothGrowth')

sample_n(ToothGrowth, size = 6)
```

Where:

* len: numeric	Tooth length
* supp: Supplement type. OJ is Orange Juice, and VC is ascorbic acid (a form 
of Vitamin C)
* dose: Dose in milligrams/day

There are 30 observations for each of the supplements, as well as 20 for each 
of the doses. This allows you to analyze the distribution of the data using 
the histogram below (Figure 1).

The graph shows that there is not a big difference between supplies when the 
dose is equal to 2mg / day. For the other doses, however, greater growth is 
noticeable when the supply is orange juice. Even so, the data show that there 
is an expected trend that increasing the dose leads to greater length 
of odontoblasts.

\newpage

```{r, echo=FALSE,warning=FALSE, message=FALSE, fig.align='center', fig.height=4, fig.width=5, fig.cap='Tooth growth (in odontoblasts) for different doses and supplements of Vitamin C'}
library(ggplot2)

ToothGrowth <- ToothGrowth %>%
  mutate(supp = if_else(supp == 'OJ', 
                        true = 'Orange Juice',
                        false = 'Ascorbic Acid'))

ggplot(ToothGrowth) +
  geom_bar(aes(x = as.factor(dose), y = len), stat = 'identity') +
  facet_wrap(~supp) +
  labs(x = 'Dose of Vitamin C (mg/day)',
       y = 'Length of odontoblasts (sum)')
```
# Inference
To begin, let's analyze the length distribution of odontoblasts looking only at
supplements. The blue lines in figure 2 represent the 99% confidence interval. 
Mean values and confidence intervals are also reported.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.height=5, fig.width=10, fig.cap='Length distribution of odontoblasts for orange juice supplement (orange line) and Ascorbic Acid (red line)'}
library(gridExtra)

vc <- ToothGrowth %>%
  filter(supp == 'Ascorbic Acid')

oj <- ToothGrowth %>%
  filter(supp == 'Orange Juice') 

stats_oj <- summary(oj$len)
stats_vc <- summary(vc$len)

conf_oj <- mean(oj$len) + c(-1, 1) * qnorm(0.99) * sd(oj$len) / sqrt(length(oj))
conf_vc <- mean(vc$len) + c(-1, 1) * qnorm(0.99) * sd(vc$len) / sqrt(length(oj))

plot_oj <- ggplot() + 
  geom_density(aes(len), oj, color = 'orange') + 
  geom_vline(xintercept = conf_oj, color = 'blue', linetype ='dashed')

plot_vc <-ggplot() +
  geom_density(aes(len), vc, color = 'red') + 
  geom_vline(xintercept = conf_vc, color = 'blue', linetype ='dashed')


grid.arrange(plot_oj, plot_vc, nrow=1)
```

```{r echo=FALSE}
message("Average length for Orange Juice: ", round(mean(oj$len), digits = 2),
        "\nAverage length for Ascorbic Acid: ", round(mean(vc$len, digits = 2)),
        "\n99% confidence interval for Orange Juice: ", round(conf_oj[1], digits = 2),
        " - ", round(conf_oj[2], digits = 2),
        "\n99% confidence interval for Ascorbic Acid: ", round(conf_vc[1], digits = 2), 
        " - ", round(conf_vc[2], digits = 2))
```

As there is not much data (20 for each distribution), it is not possible to 
assume and generalize a normality, however, there is a large concentration of 
data close to the average of the distributions and few values beyond the level 
of significance, indicating a possible trend. Still, lower values of length are 
more common in the distribution of orange juice, considering its average and, 
mainly, the lower limit of the confidence interval. The distribution for 
Ascorbic Acid contrasts with the former for the higher density (consequently, 
probability) of values close to a high average.

To analyze the influence of the dose regardless of the method used, we 
will use hypothesis tests. Looking at the histogram in figure 1, we can see 
that there is a tendency for a larger dose to positively influence the length 
of the odontoblasts, so we are assuming a null hypothesis that there is no 
influence on the dose, against an alternative hypothesis that the length of 
the odontoblasts the 2mg dose is higher than the other two.

The other hypothesis test also comes from the analysis of the histogram in 
figure 1. We can see that there does not seem to be a significant difference 
between the methods for assessing the length of odontoblasts when the dose is 
equal to 2. We will test with a null hypothesis that there is no significant 
difference, against one of which there is a significant difference (greater 
or lesser). The p-values have been rounded to the fifth decimal place.

```{r echo=FALSE}

dose_half <- ToothGrowth %>%
  filter(dose == 0.5) 
dose_one <- ToothGrowth %>%
  filter(dose == 1.0) 
dose_two <- ToothGrowth %>%
  filter(dose == 2.0) 


p_value_two_half <- t.test(dose_two$len, dose_half$len, 
       alternative = 'greater')$p.value
p_value_two_one <- t.test(dose_two$len, dose_one$len, 
       alternative = 'greater')$p.value
p_value_oj_vc <- t.test(oj$len, vc$len)$p.value

message('P-value for 0.5mg dose is more effective than 2mg dose: ', 
        round(p_value_two_half, digits = 5),
        '\nP-value for 1mg dose is more effective than 2mg dose: ',
        format(round(p_value_two_one, digits = 5), scientific = FALSE),
        '\nP-value for any significant difference between the two methods at 
        a dose of 2mg: ',
        round(p_value_oj_vc, digits = 5))

```
The p-values for the first two T tests resulted in values very close to 
zero, demonstrating that, for our data, a dose of 2mg / day is indeed more 
effective for the growth of odontoblasts compared to the other two.

On the other hand, the p-value of the last T test is not as close to zero as 
the other two (0.06). As there are approximately 6% of the data varying in the 
t-distribution tails, there will be an even greater need to measure the result 
with the level of significance $\alpha$ determined. If $\alpha$ = 0.05, we reject 
the null hypothesis (that there is no significant variation between the 
treatment methods for 2mg / day), but if we choose a $\alpha$ = 0.1, this 
null hypothesis cannot be refuted. Since we don't have a lot of data, it is 
acceptable to choose a higher level of significance instead of rejecting the 
hypothesis right away. To reduce $\alpha$, more data will be needed.

# Conclusions

As was analyzed in the previous item, despite the small amount of data, these 
seem to tend to a normal one within their confidence intervals. The 
distribution for orange juices has a distribution center close to 20, while 
the distribution center for Ascorbic Acid is facing 17 (length of odontoblasts).
The 99% confidence intervals show that there is a greater proportion of low 
length values for the distribution of Ascorbic Acid.

The hypothesis tests showed that the dose of 2mg/day is more effective and 
that it is very likely (less than 10% probability given the data) that the 
supplement influences the length of odontoblasts for a dose equal to 2mg/day

\newpage
# Appendix

```{r, message=FALSE, warning=FALSE, eval=FALSE, fig.cap='Code to load data'}
# Chunk code 1: getting the data
library('dplyr')

data('ToothGrowth')

dim(ToothGrowth) # 60 rows X 3 columns

sample_n(ToothGrowth, size = 6) # sample of six tuples

?ToothGrowth # about the data
```

```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.show='hide'}
# Figure 1 code
library(ggplot2)

table(ToothGrowth$supp) # both supplements have 30 observations
table(ToothGrowth$dose) # 20 observations for each of the three doses


# Changing OJ and VC to orange juice and ascorbic acid
ToothGrowth <- ToothGrowth %>%
  mutate(supp = if_else(supp == 'OJ', 
                        true = 'Orange Juice',
                        false = 'Ascorbic Acid'))


#plot
ggplot(ToothGrowth) +
  geom_bar(aes(x = as.factor(dose), y = len), stat = 'identity') +
  facet_wrap(~supp) +
  labs(x = 'Level of Vitamin C (mg/day)',
       y = 'Length of odontoblasts (sum)')

```

```{r warning=FALSE, message=FALSE, fig.show='hide'}
# Figure 2 code
library(gridExtra)

vc <- ToothGrowth %>%
  filter(supp == 'Ascorbic Acid')

oj <- ToothGrowth %>%
  filter(supp == 'Orange Juice') 

stats_oj <- summary(oj$len)
stats_vc <- summary(vc$len)

conf_oj <- mean(oj$len) + c(-1, 1) * qnorm(0.99) * sd(oj$len) / sqrt(length(oj))
conf_vc <- mean(vc$len) + c(-1, 1) * qnorm(0.99) * sd(vc$len) / sqrt(length(oj))

plot_oj <- ggplot() + 
  geom_density(aes(len), oj, color = 'orange') + 
  geom_vline(xintercept = conf_oj, color = 'blue', linetype ='dashed')

plot_vc <-ggplot() +
  geom_density(aes(len), vc, color = 'red') + 
  geom_vline(xintercept = conf_vc, color = 'blue', linetype ='dashed')


grid.arrange(plot_oj, plot_vc, nrow=1)
```

```{r eval=FALSE}
# Chunk code 2: Statistics
message("Average length for Orange Juice: ", round(mean(oj$len), digits = 2),
        "\nAverage length for Ascorbic Acid: ", round(mean(vc$len, digits = 2)),
        "\n99% confidence interval for Orange Juice: ", round(conf_oj[1], digits = 2),
        " - ", round(conf_oj[2], digits = 2),
        "\n99% confidence interval for Ascorbic Acid: ", round(conf_vc[1], digits = 2), 
        " - ", round(conf_vc[2], digits = 2))
```

```{r eval=FALSE}
#Chunck code 3: P-values

dose_half <- ToothGrowth %>%
  filter(dose == 0.5) 
dose_one <- ToothGrowth %>%
  filter(dose == 1.0) 
dose_two <- ToothGrowth %>%
  filter(dose == 2.0) 


p_value_two_half <- t.test(dose_two$len, dose_half$len, 
       alternative = 'greater')$p.value
p_value_two_one <- t.test(dose_two$len, dose_one$len, 
       alternative = 'greater')$p.value
p_value_oj_vc <- t.test(oj$len, vc$len)$p.value

message('P-value for 0.5mg dose is more effective than 2mg dose: ', 
        round(p_value_two_half, digits = 5),
        '\nP-value for 1mg dose is more effective than 2mg dose: ',
        format(round(p_value_two_one, digits = 5), scientific = FALSE),
        '\nP-value for any significant difference between the two methods at 
        a dose of 2mg: ',
        round(p_value_oj_vc, digits = 5))
```
