---
title: "Statistical Inference"
author: "MATIAS, V. A."
date: "07/04/2021"
output: pdf_document
subtitle: 'Part 2: Basic Inferential Data Analysis about tooth growth '
---

# Overview 
The second part of the project analyzes the dataset ToothGrowth and does 
some exploratory data analysis, and applies basic statistical inference to 
bring some information.

# Data
The dataset has 60 rows and 3 columns (appendix, chunk code 1). 
Each line has the length of the odontoblasts in the study 
(cell responsible for the growth of the teeth), the supplement used and the 
size of the dose (0.5, 1 or 2mg / day). The study was carried out on guinea 
pigs. Let's look at a random sample of 6 tuples.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library('dplyr')
data('ToothGrowth')

sample_n(ToothGrowth, size = 6)
```

Where:

* len: numeric	Tooth length
* supp: Supplement type. OJ is Orange Juice, and VC is ascorbic acid (a form 
of Vitamin C)
* dose: Dose in milligrams/day

There are 30 observations for each of the supplements, as well as 20 for each 
of the doses. This allows you to analyze the distribution of the data using 
the histogram below (Figure 1).

The graph shows that there is not a big difference between supplies when the 
dose is equal to 2mg / day. For the other doses, however, greater growth is 
noticeable when the supply is orange juice. Even so, the data show that there 
is an expected trend that increasing the dose leads to greater length 
of odontoblasts.

\newpage

```{r, echo=FALSE,warning=FALSE, message=FALSE, fig.align='center', fig.height=4, fig.width=5, fig.cap='Tooth growth (in odontoblasts) for different doses and supplements of Vitamin C'}
library(ggplot2)

ToothGrowth <- ToothGrowth %>%
  mutate(supp = if_else(supp == 'OJ', 
                        true = 'Orange Juice',
                        false = 'Ascorbic Acid'))

ggplot(ToothGrowth) +
  geom_bar(aes(x = as.factor(dose), y = len), stat = 'identity') +
  facet_wrap(~supp) +
  labs(x = 'Dose of Vitamin C (mg/day)',
       y = 'Length of odontoblasts (sum)')
```
# Inference
To begin, let's analyze the length distribution of odontoblasts looking only at
supplements. The blue lines in figure 2 represent the 99% confidence interval. 
Mean values and confidence intervals are also reported.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.height=5, fig.width=10, fig.cap='Length distribution of odontoblasts for orange juice supplement (orange line) and Ascorbic Acid (red line)'}
library(gridExtra)

vc <- ToothGrowth %>%
  filter(supp == 'Ascorbic Acid')

oj <- ToothGrowth %>%
  filter(supp == 'Orange Juice') 

stats_oj <- summary(oj$len)
stats_vc <- summary(vc$len)

conf_oj <- mean(oj$len) + c(-1, 1) * qnorm(0.99) * sd(oj$len) / sqrt(length(oj))
conf_vc <- mean(vc$len) + c(-1, 1) * qnorm(0.99) * sd(vc$len) / sqrt(length(oj))

plot_oj <- ggplot() + 
  geom_density(aes(len), oj, color = 'orange') + 
  geom_vline(xintercept = conf_oj, color = 'blue', linetype ='dashed')

plot_vc <-ggplot() +
  geom_density(aes(len), vc, color = 'red') + 
  geom_vline(xintercept = conf_vc, color = 'blue', linetype ='dashed')


grid.arrange(plot_oj, plot_vc, nrow=1)
```

```{r echo=FALSE}
message("Average length for Orange Juice: ", round(mean(oj$len), digits = 2),
        "\nAverage length for Ascorbic Acid: ", round(mean(vc$len, digits = 2)),
        "\n99% confidence interval for Orange Juice: ", round(conf_oj[1], digits = 2),
        " - ", round(conf_oj[2], digits = 2),
        "\n99% confidence interval for Ascorbic Acid: ", round(conf_vc[1], digits = 2), 
        " - ", round(conf_vc[2], digits = 2))
```

As there is not much data (20 for each distribution), it is not possible to 
assume and generalize a normality, however, there is a large concentration of 
data close to the average of the distributions and few values beyond the level 
of significance, indicating a possible trend. Still, lower values of length are 
more common in the distribution of orange juice, considering its average and, 
mainly, the lower limit of the confidence interval. The distribution for 
Ascorbic Acid contrasts with the former for the higher density (consequently, 
probability) of values close to a high average.

```{r}

t.test(oj$len, vc$len, conf.level = 0.99, alternative = 'greater')


```

# Conclusions


# Appendix

```{r, message=FALSE, warning=FALSE, eval=FALSE, fig.cap='Code to load data'}
# Chunk code 1: getting the data
library('dplyr')

data('ToothGrowth')

dim(ToothGrowth) # 60 rows X 3 columns

sample_n(ToothGrowth, size = 6) # sample of six tuples

?ToothGrowth # about the data
```

```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.show='hide'}
# Figure 1 code
library(ggplot2)

table(ToothGrowth$supp) # both supplements have 30 observations
table(ToothGrowth$dose) # 20 observations for each of the three doses


# Changing OJ and VC to orange juice and ascorbic acid
ToothGrowth <- ToothGrowth %>%
  mutate(supp = if_else(supp == 'OJ', 
                        true = 'Orange Juice',
                        false = 'Ascorbic Acid'))


#plot
ggplot(ToothGrowth) +
  geom_bar(aes(x = as.factor(dose), y = len), stat = 'identity') +
  facet_wrap(~supp) +
  labs(x = 'Level of Vitamin C (mg/day)',
       y = 'Length of odontoblasts (sum)')

```

```{r warning=FALSE, message=FALSE, fig.show='hide'}
# Figure 2 code
library(gridExtra)

vc <- ToothGrowth %>%
  filter(supp == 'Ascorbic Acid')

oj <- ToothGrowth %>%
  filter(supp == 'Orange Juice') 

stats_oj <- summary(oj$len)
stats_vc <- summary(vc$len)

conf_oj <- mean(oj$len) + c(-1, 1) * qnorm(0.99) * sd(oj$len) / sqrt(length(oj))
conf_vc <- mean(vc$len) + c(-1, 1) * qnorm(0.99) * sd(vc$len) / sqrt(length(oj))

plot_oj <- ggplot() + 
  geom_density(aes(len), oj, color = 'orange') + 
  geom_vline(xintercept = conf_oj, color = 'blue', linetype ='dashed')

plot_vc <-ggplot() +
  geom_density(aes(len), vc, color = 'red') + 
  geom_vline(xintercept = conf_vc, color = 'blue', linetype ='dashed')


grid.arrange(plot_oj, plot_vc, nrow=1)
```

```{r eval=FALSE}
# Chunk code 2: Statistics
message("Average length for Orange Juice: ", round(mean(oj$len), digits = 2),
        "\nAverage length for Ascorbic Acid: ", round(mean(vc$len, digits = 2)),
        "\n99% confidence interval for Orange Juice: ", round(conf_oj[1], digits = 2),
        " - ", round(conf_oj[2], digits = 2),
        "\n99% confidence interval for Ascorbic Acid: ", round(conf_vc[1], digits = 2), 
        " - ", round(conf_vc[2], digits = 2))
```
