---
title: Analysis of severe weather events in the united states between the years 1950
  and 2011
author: "MATIAS, V. A."
date: "28/03/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Synopsis
<p> This is the final project of the 
[Reproducible Research Course](https://www.coursera.org/learn/reproducible-research) 
offered by Johns Hopkins University on the 
[Coursera](https://www.coursera.org/) platform.
<p> This paper will present an analysis of climate events based on the U.S. 
National Oceanic and Atmospheric Administration's (NOAA) storm database.
<p> The programming language R will be used as the basis for all the analysis 
to be presented, this document being developed with R markdown (Rmd).

## 1.1. Data
<p> The available data are in a compressed CSV file for type .bz2, 
resulting in a 42Mb file.

<p> Data: [Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)
<p> Documentation: [National Weather Service Storm Data Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
<p> FAQ: [National Climatic Data Center Storm Events FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)

# 2. Data Processing
<p> This section covers the techniques used in the collection and processing of data

## 2.1. Required libraries
<p> Before starting the analysis, some packages will be needed to facilitate 
development:

* dplyr: To use the function mapvalues, a good way to use the idea of hashmap in R
* dplyr: To facilitate data manipulation
* ggplot2: To create the graphics 
* gridExtra: To create a view with more than one panel
* data.table: To use the _fread()_ function, being faster to read a CSV than 
 the _read.csv()_ function
* R.utils: To allow the fread function to read a CSV without having to unzip 
a bz2 file
```{r warning=FALSE, message=FALSE}
library(plyr) # Must be called before dplyr
library(dplyr)
library(ggplot2)
library(gridExtra)
library(data.table)
library(R.utils)
```

## 2.2. Geting the data
<p> The project requires that the compressed CSV files be in the directory 
where the analysis will be performed. This step can be performed by the user 
himself, downloading the file and placing it in the analysis directory, or 
by the code below.
```{r eval=FALSE}
# Optional step
data_url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(data_url, destfile = "storm_data.csv.bz2")
```

## 2.3. Loading the data
<p> Now the downloaded file with the name _storm_data.csv.bz2_ will be stored in
memory. Note that this is a large file when uncompressed, requiring almost 500Mb
to allocate it in RAM.

<p> To avoid storing 500Mb in RAM, we will only select the columns that will 
be used in the analysis. They are: PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP, 
INJURIES, FATALITIES, EVTYPE. This approach results in just under 50Mb to be 
stored in memory.
```{r cache=TRUE}
weather <- fread("storm_data.csv.bz2") %>%
  as_tibble() %>%
  select(PROPDMG, 
         PROPDMGEXP, 
         CROPDMG, 
         CROPDMGEXP, 
         INJURIES, 
         FATALITIES, 
         EVTYPE) %>%
  mutate(EVTYPE = as.factor(EVTYPE))
object.size(weather)
```

# 3. Results
<p> The analysis of the data and its consequent results comes from the 
questions proposed by the Assignment.

## 3.1. Harmful events
<p> The first analysis to be made is: given the different types of recorded 
climatic events, which are the most harmful to the population health? The code 
below realized some transformations in the original data to get the events 
with more frequencies of deaths and injuries. At the end is created one 
figure composed of two histograms.
```{r}
#Event types with most injured people
injuries_evtype <- weather %>%
  group_by(EVTYPE) %>%
  summarise(sum_injuries = sum(INJURIES), .groups = "drop") %>%
  arrange(desc(sum_injuries)) %>% 
  head()
injuries_evtype

#Event types with most fatalities
fatal_evtype <- weather %>%
  group_by(EVTYPE) %>%
  summarise(sum_fatalities = sum(FATALITIES), .groups = "drop") %>%
  arrange(desc(sum_fatalities)) %>% 
  head()
fatal_evtype

#Bar plot of the events with the most injuries
injuries_plot <- ggplot(injuries_evtype, aes(x = EVTYPE, y = sum_injuries)) + 
  geom_bar(stat = "identity") +
  labs(x = "Event Type",
       y = "Injured people",
       title = "Most injury-causing events") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

#Bar plot of the events with the most deaths
fatal_plot <- ggplot(fatal_evtype, aes(x = EVTYPE, y = sum_fatalities)) + 
  geom_bar(stat = "identity") +
  labs(x = "Event Type",
       y = "Deaths",
       title = "Most fatal events") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

#Figure of the most harmful events
gridExtra::grid.arrange(injuries_plot, fatal_plot, 
                        nrow = 1, 
                        top = "Figure 1 - Most harmful events")
```
<p> As we can see, the six events with more causes of injuries also are the 
same events with more causes of deaths.  Tornadoes are the most harmful 
events pointed out by the data.

## 3.2. Economic consequences
<p> The second question is to analyze which climatic events have caused the 
most economic damage.

<p> From the review of the database documentation and the analysis of the 
variables in the CSV file, it is possible to identify that the PROPDMG 
attribute refers to the damage caused on properties, just as CROPDMG refers 
to the damage caused on crops.

<p> The documentation refers to factors K, M and B as well. These factors mean 
thousands, millions and billions, respectively. The PROPDMGEXP and CROPDMGEXP 
variables store these references for the loss values. There are still other 
values not specified in the documentation but present in these columns (such 
as numerical values and letters with no relation, such as "H"), these values 
were disregarded because there was no way to measure them.
```{r warning=FALSE}
# The two vectors below will be used in the function mapvalues, working like 
# a numeric reference for each character (K = 10^3, M = 10^6 and B = 10^9)
units_known <- c("K", "M", "B")
ref_known <- c(10^3, 10^6, 10^9)

#Creating the data with the value in dollars (2011) of damage in crops and properties
dmg <- weather %>%
  filter(PROPDMGEXP %in% units_known | CROPDMGEXP %in% units_known) %>%
  mutate(PROPDMGEXP = mapvalues(PROPDMGEXP, from = units_known, to = ref_known),
         CROPDMGEXP = mapvalues(CROPDMGEXP, from = units_known, to = ref_known),
         PROPDMG = PROPDMG * as.numeric(PROPDMGEXP),
         CROPDMG = CROPDMG * as.numeric(CROPDMGEXP)) %>%
  group_by(EVTYPE) %>%
  summarise(sum_propdmg = sum(PROPDMG, na.rm = TRUE),
            sum_cropdmg = sum(CROPDMG, na.rm = TRUE),
            total_dmg = sum_propdmg + sum_cropdmg,
            .groups = "drop") %>% 
  arrange(desc(total_dmg))
dmg

#Making the plot (here we will use just the 6 events with most damage)
ggplot(head(dmg), aes(x = EVTYPE, y = total_dmg/10^9)) + 
  geom_bar(stat = "identity") + 
  labs(x = "Event Type",
       y = "Damage (billions of dollars) ",
       title = "Figure 2 -Types of events with the greatest economic consequences") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))
```
<p> Although tornadoes have the greatest responsibility for the damage to the 
health of the population, they account for just over one third of the greatest 
cause of economic damage: Floods. It is worth noting that, even so, tornadoes 
are the third biggest cause of economic losses.


