---
title: "Analysis of severe weather events in the united states between the years 1950 and 2011"
author: "MATIAS, V. A."
date: "28/03/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Synopsis
<p> This is the final project of the 
[Reproducible Research Course](https://www.coursera.org/learn/reproducible-research) 
offered by Johns Hopkins University on the 
[Coursera](https://www.coursera.org/) platform.
<p> This paper will present an analysis of climate events based on the U.S. 
National Oceanic and Atmospheric Administration's (NOAA) storm database.
<p> The programming language R will be used as the basis for all the analysis 
to be presented, this document being developed with R markdown (Rmd).

## 1.1 Data
<p> The available data are in a compressed CSV file for type .bz2, 
resulting in a 42Mb file.

<p> Data: [Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)
<p> Documentation: [National Weather Service Storm Data Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
<p> FAQ: [National Climatic Data Center Storm Events FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)

# 2. Data Processing
<p> This section covers the techniques used in the collection and processing of data

## 2.1 Required libraries
<p> Before starting the analysis, some packages will be needed to facilitate development:

* dplyr: To facilitate data manipulation
* ggplot2: To create the graphics 
```{r warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
```

## 2.2 Geting the data
<p> The project requires that the compressed CSV files be in the directory 
where the analysis will be performed. This step can be performed by the user 
himself, downloading the file and placing it in the analysis directory, or 
by the code below.
```{r eval=FALSE}
# Optional step
data_url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(data_url, destfile = "storm_data.csv.bz2")
```

## 2.3 Loading the data
<p> Now the downloaded file with the name _storm_data.csv.bz2_ will be stored in
memory. Note that this is a large file when uncompressed, requiring almost 500Mb
to allocate it in RAM.
```{r cache=TRUE}
weather <- read.csv("storm_data.csv.bz2")
object.size(weather)
```


